A weighted data representation based approach for manifold learning

### 2018-11-01 DimReduction/Lle/Documentation/Article_lle.tex
Adding first part of introduction to Lle documentation

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\begin{abstract}
    We implemented the weighted data representation based in KNN and minimization
    of the euclidean distance between representation. Finally the local embedding
    is tested on 3-dimensional manifolds.
\end{abstract}

@@

@@

\subsection{First optimization problem}

@@

    \subsection{Case for an arbitrary $i$}

    define

@@

    then

@@

@@

    define

@@

    Now $x_i$ can be defined as

@@

@@

    Define a matrix $G$ as

@@

    Redefine the problem as

@@

    To solve this problem the Lagrange multipliers are used. First redefine the
    restriction as

@@

    Now

    \begin{equation*}
        \begin{aligned}
            \mathcal{L}(W_i, \lambda) &= W_i^t G W_i - \lambda (e^t W_i - 1)\\
            \frac{\delta \mathcal{L}}{\delta W_i} &= 2 G W_i - \lambda e
        \end{aligned}
    \end{equation*}

    \begin{equation*}
        \begin{aligned}
            2 G W_i - \lambda e &= 0 \\
            2 G W_i &= \lambda e \\
            G W_i &= \frac{1}{2} \lambda e \\
            G^{-1} G W_i &= G^{-1} \frac{1}{2} \lambda e \\
            W_i &= \frac{1}{2} \lambda G^{-1} e \\
        \end{aligned}
    \end{equation*}

    Now if $\lambda$ is wrongly chosen, then the result would be a scaled version
    of the true value. Then, $\lambda$ can be chosen arbitrarily, and after $W_i$
    is computed a rescale procedure can be conducted in order to satisfy the
    restriction.\\

    For simplicity choose $\lambda = 2$, then

    \begin{equation*}
        \begin{aligned}
            W_i &= \frac{1}{2} \lambda G^{-1} e = G^{-1} e \\
            G W_i &= e \\
        \end{aligned}
    \end{equation*}

    So the problem is reduced to solve a linear system of equations. Now if
    after build the matrix G, it turns out that is a singular matrix the
    following computation can be used to make it invertible

    \begin{equation*}
            G = G + \sigma I
    \end{equation*}

    where sigma is a small value.

    \subsection{Second optimization problem}

    \begin{equation*}
        \begin{aligned}
            \underset{y}{\text{min}} \quad \sum_{i=1}^n & \lVert y_i - \sum_{j=1}^n w_{ij} y_j \rVert^2 \\
            & YY^t = I
        \end{aligned}
    \end{equation*}

@@

    Now, define

    \[
        W = 
        \begin{bmatrix}
            |   &  |  &       & |   \\
            w_1 & w_2 & \dots & w_n \\
            |   &  |  &       & |
        \end{bmatrix}_{n \times n}
        where
        \hspace{0.5cm}
        W_i = 
        \begin{bmatrix}
            0 \\
            \vdots \\ 
            w_1 \\
            \vdots \\ 
            w_k \\
            \vdots \\ 
            0 \\
        \end{bmatrix}_{n \times 1}
    \]

    Where $W_i$ is a vector that contains the value of the weight that correspond
    to the $j$ neighbor of $x_i$ in the corresponding position, and 0 anywhere
    else. This means that

    \begin{equation*}
        X W_i = \sum_{j=1}^k w_{ij} x_j
    \end{equation*}

    Rewrite the problem as

    \begin{equation*}
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - \sum_{j=1}^n w_{ij} y_j \rVert^2
            =
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - Y W_i \rVert^2
    \end{equation*}

    Also define

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{babelbib}

@@

\begin{abstract}
    We present a detailed implementation of the local linear embeding
    algorithm from both a theorical and practical approximation. In this work the
    dimensionality reduction is achieved by a two step procedure. First a
    weighted data representation based in $K$ nearest neighbors is calculated, and
    then a minimization of distance between the data representation and a low
    dimensionality configuration is conducted. Finally the resulting algorithm
    is tested on 3-dimensional manifolds.
\end{abstract}

@@

Exploratory data analysis and visualization are key issues in a lot of areas
in science. Gather knowledge about the structure and caracteristics of data
can lead to great improves in the process of solving problems in science. But a
good analysis of multidimensinal data can be really hard to achive, and the
consecuences of mistaken assumptions about the data properties can produce
dysfunctional solutions.

\vspace{0.25cm}

An approximation to solve this problems are dimensionality reduction techniques.
These methods provide a way to discover compact representations of
multidimensional datasets in a low dimensional space. This approaches allow the
analysis process to be carried out in an easier way and help to reduce the
probability of getting erroneous results.

\vspace{0.25cm}

However, there are still some issues related to the usage of these methods.
Assumptions like the use of euclidean distances as a measure in the
multidimensional space implies that the data is homogeneous or isotropic, and
this is really often not the case in many real world datasets
\cite{homogeneous1}, \cite{homogeneous2}. Because of this, the usage of other
distance metrics is a good option to increase the performance of these approches
\cite{dist1}, \cite{dist2}.

\vspace{0.25cm}

An appropriate processing method for dimentionality reduction in the case of
a nonlinear approach can be manifold learning techniques \cite{manifold1},
\cite{manifold2}. This algorithms seek to generate nonlinear dimensionality
reductions by exploiting

% missing part
\cite{lle}

In this work the linear local embeding algorithm is used to perform a
dimensionality reduction of a multidimentional dataset by a manifold learning
technique. The first part of the work will carafully describe each part of the
mathemathical backgound of the method, then the implementation details of the
algorithm are shown to prepare a results section, and finally some conclusions
are exposed.

@@

    As mention before, the problem that we are trying to solve is to find
    a low dimensional representation of points in a manifold that is in a high
    dimensional space. In order to achieve this, two steps are considered. The
    first one is to find a way to represent each point in the original dataset as
    a linear combination of its $K$ nearest neighbors. This means that the
    objective is to find a set of weights that minimize the following
    expression

@@

    Where $N(x_i)_j$ represents the $j$-th neighbor of the point $x_i$. The
    restriction is imposed to get a non trivial solution and ensure that
    the construction of the points is convex.\\

    This problem can be solved by optimizing an arbitrary case $i$. The first
    step is to define the neighborhood matrix of $x_i$, a vector that will
    conatin the weights of the combination and a special matrix $e$ that will
    facilitate the process later.

@@

    Whith these tools the linear combinations of neighbors can be expressed as

@@

    Hence the problem can be rewritten as follows

@@

    For the next step the objective is to represent $x_i$ as a vector operation.
    First a matrix in which each column is the $x_i$ point is constructed with
    the help of the $e$ vector

@@

    Then using the fact that $W_i$ is used to construct a convex representation
    of the desired point, $x_i$ can be computed as

@@

    Now the problem can be redefined in terms of matrix and vectors operations

@@

    The next step is to define a matrix to simplify the objective function

@@

    With this matrix the problem can be rewritten as follows

@@

    An option to solve this problem are Lagrange multipliers. In order to use
    this method the restriction is redefined

@@

    Then a new objective function can be constructed

    \begin{equation*}
            \mathcal{L}(W_i, \lambda) = W_i^t G W_i - \lambda (e^t W_i - 1)
    \end{equation*}

    In order to optimize this function its derivative is calculated and then is
    set to 0

    \begin{equation*}
        \begin{aligned}
            \frac{\delta \mathcal{L}}{\delta W_i} &= 2 G W_i - \lambda e\\
            2 G W_i - \lambda e &= 0 \\
            2 G W_i &= \lambda e \\
            G W_i &= \frac{1}{2} \lambda e \\
            G^{-1} G W_i &= G^{-1} \frac{1}{2} \lambda e \\
            W_i &= \frac{1}{2} \lambda G^{-1} e \\
        \end{aligned}
    \end{equation*}

    At this point, if $\lambda$ is wrongly chosen, then the result would be a
    scaled version of the true value. Then, $\lambda$ can be chosen
    arbitrarily. Then, after $W_i$ is computed a rescale procedure can be
    conducted in order to satisfy the restriction.\\

    A good choice for the variable is $\lambda = 2$. This leads to the
    cancellation of the fraction and reduces the problem to find the solution of
    a linear system of equations

    \begin{equation*}
        \begin{aligned}
            W_i = \frac{1}{2} \lambda G^{-1} & e = G^{-1} e \\
            G W_i &= e \\
        \end{aligned}
    \end{equation*}

    This implies that matrix $G$ must be invertible, and there is no warranty
    of that. So, if that is the case, a solution is to recalculate $G$ as

    \begin{equation*}
        G \leftarrow G + \sigma I
    \end{equation*}

    where $\sigma$ is a small value.\\

    This procedure is used to find all the weight values. Then this weights are
    used to define a configuration of points in a low dimensional space that
    is as similar as possible to the original dataset. The objective in this
    point of the problem is minimize the following expression

    \begin{equation*}
        \begin{aligned}
            \underset{y}{\text{min}} \quad \sum_{i=1}^n & \lVert y_i - \sum_{j=1}^n w_{ij} N(y_i)_j \rVert^2 \\
            & YY^t = I
        \end{aligned}
    \end{equation*}

@@

    To solve this problem define

    \[
        W =
        \begin{bmatrix}
            |   &  |  &       & |   \\
            W_1 & W_2 & \dots & W_n \\
            |   &  |  &       & |
        \end{bmatrix}_{n \times n}
        where
        \hspace{0.5cm}
        W_i =
        \begin{bmatrix}
            0 \\
            w_1 \\
            \vdots \\
            w_k \\
            0 \\
        \end{bmatrix}_{n \times 1}
    \]

    Where $W_i$ is a vector that contains the value of the weight that correspond
    to the $j$ neighbor of $x_i$ in the corresponding position, and 0 anywhere
    else.\\

    Using this matrix the linear combinations of points can be calculated as

    \begin{equation*}
        \sum_{j=1}^k w_{ij} N(y_i)_j = Y W_i
    \end{equation*}

    Which means that the problem can be rewrited as

    \begin{equation*}
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - \sum_{j=1}^n w_{ij} N(y_i)_j \rVert^2
            =
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - Y W_i \rVert^2
    \end{equation*}

    Now,

@@

\bibliographystyle{unsrt}
\bibliography{Article_lle}
